---
title: "Flight Data Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(plotly)
```



```{r}
rawDB <- read.csv("FinalFlightsNumeric_with_Aircraft_Info.csv")
```


```{r}
mainDB <- rawDB %>% 
  select(UNIQUE_CARRIER, 
         TAIL_NUM,
         ORIGIN_AIRPORT_ID, 
         DEST_AIRPORT_ID,        
         DISTANCE, 
         CRS_ELAPSED_TIME,
         ARR_DELAY,
         CARRIER_DELAY,     
         WEATHER_DELAY,      
         NAS_DELAY,     
         SECURITY_DELAY,   
         LATE_AIRCRAFT_DELAY,           
         CARRIER_NAME,   
         ORIGIN_AIRPORT,   
         DESTINATION_AIRPORT, 
         CARRIER_CODE,  
         YEAR.MFR,    
         MFR.MDL.CODE,   
         MFR,      
         MODEL,       
         NO.ENG,     
         NO.SEATS)
```

```{r}
aircraft_types <- unique(mainDB$MODEL)
write.csv(aircraft_types, "aircraft_types.csv")
```


```{r}

flight_countDB <- mainDB %>%
  group_by(MODEL) %>%
  summarize(FLIGHT_COUNT = n()) %>%
  arrange(desc(FLIGHT_COUNT))



```

#export to csv to do some manual cleanup in excel
```{r}
write.csv(flight_countDB, "aircraft_types.csv")
```


#import this new version of the csv 
```{r}
flight_countDB1 <- read.csv("aircraft_types.csv")
```

```{r}
aircraft_types_mapping <- flight_countDB1 %>% 
  select(MODEL, MODEL_GENERAL)
```



#merge the new version back to the old table

```{r}
mainDB_aircraft_types <- left_join(mainDB, aircraft_types_mapping, by = c("MODEL", "MODEL"))
```




#Get number of flights by model general
```{r}
mainDB_aircraft_types_flight_count <- mainDB_aircraft_types %>%
  group_by(MODEL_GENERAL) %>%
  summarize(FLIGHT_COUNT_SUM = n()) %>%
  arrange(desc(FLIGHT_COUNT_SUM))
```



#make vector of top aircraft
```{r}

mainDB_aircraft_types_flight_count <- mainDB_aircraft_types_flight_count %>% 
  filter(MODEL_GENERAL != "NA") %>%
  filter(MODEL_GENERAL!= "OTHER")


top_aircraft <- head(as.vector(mainDB_aircraft_types_flight_count$MODEL_GENERAL), 7)
top_aircraft

```

#filter out unpopular aircraft, and flights that dont have elapsed time info
```{r}
mainDB_aircraft_types <- mainDB_aircraft_types %>%
  filter(MODEL_GENERAL %in% top_aircraft) %>%
  filter(!is.na(CRS_ELAPSED_TIME)) %>%
  filter(!is.na(ARR_DELAY)) %>%
  filter(ORIGIN_AIRPORT == "Boston, MA: Logan International")%>%
  droplevels()

```

```{r}
mainDB_aircraft_types$ROUTE <- paste(mainDB_aircraft_types$ORIGIN_AIRPORT,"to", mainDB_aircraft_types$DESTINATION_AIRPORT)
```



#get most popular route by airplane

```{r}
dataviz_table <- mainDB_aircraft_types %>% 
  group_by(MODEL_GENERAL) %>%
  summarize(FLEET_SIZE = length(unique(TAIL_NUM)),
            
            POPULAR_ROUTE = names(which.max(table(ROUTE))),
            DESTINATION_AIRPORT = names(which.max(table(DESTINATION_AIRPORT))),
            POPULAR_CARRIER = names(which.max(table(CARRIER_NAME))),
            NUM_FLIGHTS = n(),
            AVG_NUM_FLIGHTS_PLANE = round(NUM_FLIGHTS/FLEET_SIZE,2),
            AVG_MILES_FLIGHT = round(sum(DISTANCE)/NUM_FLIGHTS,2),
            AVG_TIME_FLIGHT = round(sum(CRS_ELAPSED_TIME)/NUM_FLIGHTS,2),
            AVG_ARR_DELAY  = round(mean(ARR_DELAY),2),
            ON_TIME_PERF = round(sum((ARR_DELAY < 15))/NUM_FLIGHTS,2) ,
            AVG_SEAT = round(mean(NO.SEATS),0), 
            AVG_AGE = round(mean(2017 - YEAR.MFR),2),
            NUM_ENGINES = round(mean(NO.ENG),0)
            
            
            
            
            
            )    


```
#Save out the final data table
```{r}
write.csv(dataviz_table , "dataviz_table.csv")
```










```{r}
 hist_table <- mainDB_aircraft_types %>% 
  filter(ORIGIN_AIRPORT == "Boston, MA: Logan International")%>%
  select(MODEL_GENERAL, ARR_DELAY)
        
```

```{r}

hist_table_test <- transform(hist_table, sum = ave(ARR_DELAY, CARRIER_CODE, FUN = cumcount))
```

```{r}
hist_table$num<- ave(hist_table$ARR_DELAY, hist_table$CARRIER_CODE, FUN = seq_along)
```

```{r}
library(tidyr)

```

```{r}



hist_table_test <- hist_table %>% 
  mutate(i = row_number()) %>% 
  spread(MODEL_GENERAL, ARR_DELAY)
```

```{r}
hist_table_test$i <- NULL
```



```{r}
write.csv(hist_table, "hist_table.csv")
```

------

```{r}
bar_table <- mainDB_aircraft_types %>%
  group_by(MODEL_GENERAL) %>%
  summarize(NUM_FLIGHTS = n(),
    ON_TIME_PERF = round(sum((ARR_DELAY < 15))/NUM_FLIGHTS,3))

bar_table <- bar_table %>%
  select(MODEL_GENERAL, ON_TIME_PERF)
```


```{r}
write.csv(bar_table, "bar_table.csv")
```

------------------
#plotly



```{r}
library(plotly)

Sys.setenv("plotly_username"="chilger")
Sys.setenv("plotly_api_key"="XTYltgJFx1jAaiBAzigK")


```




#Year of manufacture of flights departing Boston

```{r}



mainDB_aircraft_types$MODEL_GENERAL <- factor(mainDB_aircraft_types$MODEL_GENERAL, levels =c("A321","737","A320","A319","717","757","CL-600"))


p <- plot_ly(mainDB_aircraft_types, y= ~YEAR.MFR, x = ~MODEL_GENERAL, color = ~MODEL_GENERAL, type = "box") 

#p <- plot_ly(ggplot2::diamonds, y = ~price, color = ~cut, type = "box")
p

# Create a shareable link to your chart
# Set up API credentials: https://plot.ly/r/getting-started

#chart_link = api_create(p, filename="box/jitter")
#chart_link
```


#Number of Seats of flights departing boston
```{r}
mainDB_aircraft_types$MODEL_GENERAL <- factor(mainDB_aircraft_types$MODEL_GENERAL, levels =c("A321","737","A320","A319","717","757","CL-600"))


p <- plot_ly(mainDB_aircraft_types, y= ~NO.SEATS, x = ~MODEL_GENERAL, color = ~MODEL_GENERAL, type = "box") 

#p <- plot_ly(ggplot2::diamonds, y = ~price, color = ~cut, type = "box")
p
```


#delay categorization

```{r}
mainDB_aircraft_types_delay <- mainDB_aircraft_types

```

```{r}
mainDB_aircraft_types_delay$DELAY_LEVEL <-cut(mainDB_aircraft_types_delay$ARR_DELAY, 
                                              breaks = c(-Inf,0, 15, 60, 180, Inf),
                                              labels = c("Early", "< 15 Min", "< 60 Min", "< 180 Min", "Fucking late"),
                                              right = FALSE)

  
  


```

#group_by plane and then delay type, summarize by % of each delay type
```{r}
mainDB_aircraft_types_delay_grouped <- mainDB_aircraft_types_delay %>% 
  group_by(MODEL_GENERAL, DELAY_LEVEL) %>%
  summarize(n = n()) %>%
  mutate(PERCENT_OCCURANCE = round(n / sum(n), 4))




```






```{r}
mainDB_aircraft_types_delay_grouped$DELAY_LEVEL <- factor(mainDB_aircraft_types_delay_grouped$DELAY_LEVEL, levels =rev(c("Fucking late", "< 180 Min","< 60 Min", "< 15 Min", "Early")))
```




```{r}


 
# Stacked Percent
ggplot(mainDB_aircraft_types_delay_grouped, aes(fill=DELAY_LEVEL, y=PERCENT_OCCURANCE, x=MODEL_GENERAL)) + 
    geom_bar( stat="identity", position="fill")
 
```


```{r}

f <- list(
  size = 18
)

x <- list(
  title = "Percentage",
  titlefont = f,
  nticks = 10,
  tickformat = "%")

y <- list(
  title = "Model of Airplane",
  titlefont = f)

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)






delay_plot <- plot_ly(mainDB_aircraft_types_delay_grouped, y = ~MODEL_GENERAL, x = ~PERCENT_OCCURANCE, color = ~DELAY_LEVEL, colors = rev(c('#d73027','#fc8d59','#fee08b','#a6d96a','#66bd63')), opacity = "0.8", type = "bar") %>%
  layout(xaxis = x,yaxis = y,bargap=0.2, barmode = 'stack', title = "Percent Occurance of Delay", titlefont = c(size = 40), width = 500, height = 700, margin = m )


delay_plot


chart_link = plotly_POST(delay_plot, filename="airline_delay")

chart_link

```


